---
title: "**01 Pokèmon Data Visualization with R**"
author: "Zinni Matteo"
date: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r reading dataset, include=FALSE}
pokemon = read.csv(file.choose(), dec = ".", quote = "\"")
pokemon = pokemon[,-1]
```

## Background
The Complete Pokemon Dataset includes `r dim(pokemon)[1]` Pokèmon and `r dim(pokemon)[2]` variables from 7 generations (1996-2019). Features refer only to Pokèmon video games and not to other products such as Pokèmon GO or trading card game. The original version is available for free download at [Kaggle](https://www.kaggle.com/rounakbanik/pokemon).

## Objective
The first project within Pokémon data science is focused on data visualization as a powerful tool to give a first overview of data for further analysis. The dataset represent an interesting opportunity to display both quantitative (continuous and discrete) and qualitative variables.  



```{r edit df, include=FALSE}
#remove japanese name column
pokemon = pokemon[,-29]

# make the the first letter capital
headers = gsub("(?<=^|_)([a-z])", "\\U\\1", colnames(pokemon), perl=TRUE)

# replace headers
colnames(pokemon) = headers

# edit some variables names to look smarter
colnames(pokemon)[27] <- "Height_m"
colnames(pokemon)[35] <- "Type_1"
colnames(pokemon)[36] <- "Type_2"

# edit generation name to display roman numeral
Generation = factor(pokemon$Generation, levels=c(1:7), labels = as.roman(c(1:7)))

# replace column
pokemon$Generation <- Generation
```


```{r check head df, echo=FALSE}
knitr::kable(pokemon)
```





